\nohyphens{\textbf{Automated~Analysis~of~Weak~Memory~Models}}

### 2018-06-15 slides/slides.tex
Consider Professor's comments. Cleanup and extend.
Improve slides.

\documentclass{beamer}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian,english]{babel}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.pathmorphing,decorations.markings}
\usetikzlibrary{shapes}
\usetikzlibrary{patterns}
\usetikzlibrary{automata}
%\usetikzlibrary{tikzmark}
%\usetikzlibrary{calc}

@@

@@

\usepackage{listings}

@@

\usepackage[backend=biber]{biblatex}
\addbibresource{../bib/ms-thesis.bib}
%\newcommand{\tikzmark}[1]{\tikz[remember picture] \node[coordinate] (#1) {#1};}

\usepackage{array}
\usepackage{xcolor}

@@

% There are many different themes available for Beamer. A comprehensive
% list with examples is given here:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
% You can uncomment the themes below if you would like to use a different
% one:
%\usetheme[school={SCI}]{Aalto}
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{default}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

%\beamertemplatenavigationsymbolsempty  %no navigation bar
%\setbeamertemplate{footline}[frame number]
%\setbeamertemplate{headline}{}
\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
}

\setbeamertemplate{frametitle}[default][left]

%\titlegraphic{\includegraphics[height=1cm,width=2cm]{logo1}}
%\titlegraphicii{\includegraphics[height=1cm,width=2cm]{logo2}}

\title{\nohyphens{\textbf{Automated~Analysis~of~Weak~Memory~Models}}}

@@

\institute % (optional, but mostly needed)
{
  \inst{1}%
  Department of Computer Science, \\
  School of Science, \\
  \textbf{Aalto University} (Espoo, Finland)%
  \and
  \inst{2}%
  Faculty of Information Security \\
  and Computer Technologies, \\
  \textbf{ITMO University} (Saint Petersburg, Russia)%
}

@@

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
\AtBeginSection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection]%,currentsubsection]
  \end{frame}
}

\definecolor{ltgrey}{rgb}{0.85,0.85,0.85}
\definecolor{dkblue}{rgb} {0,   0.1, 0.6}
\definecolor{dkred} {rgb} {0.6, 0.1, 0}
\renewcommand{\r}[1]{\ensuremath{\textcolor{dkred}{#1}}}
\renewcommand{\b}[1]{\ensuremath{\textcolor{dkblue}{#1}}}

% Let's get started
\begin{document}

@@

\begin{frame}{Outline}
  \tableofcontents
\end{frame}

\section{Introduction}

\subsection{The problem}

\begin{frame}{Problem statement (Цель работы)}
To rework the proof-of-concept memory model-aware analysis tool \texttt{Porthos}~\cite{Porthos17a} by:
\begin{itemize}
\item extending the C-like input language,
\item revising its architecture and
\item re-implementing the tool in order to enhance performance, extensibility, reliability and maintainability
\end{itemize}
\end{frame}

\begin{frame}{Task specification (Задачи работы)}
\begin{itemize}
\item Study the general framework for memory model-aware analysis of concurrent programs~\cite{alglave2010shared};
\item Review existing tools for memory model-aware analysis;
\item Investigate existing architecture of \texttt{Porthos}, its strengths and weaknesses;
\item Design a new architecture for \texttt{PorthosC} that \textit{allow to} easily support the C input language, be robust, transparent, efficient and extensible.
\end{itemize}
\end{frame}

\subsection{The background: weak memory model-aware analysis}

\begin{frame}{Verification of concurrent software} {Example: Write-write reordering (compiler  relaxations)}
%1) mckenney  http://www.rdrop.com/users/paulmck/scalability/paper/whymb.2010.06.07c.pdf
%2) http://preshing.com/20120515/memory-reordering-caught-in-the-act/

\begin{table}
\centering
\ttfamily
\begin{tabular}{ |>{\color{dkblue}}c | >{\color{dkred}}c| }
\hline
\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P & Q \\ \hline
$p_0$ : $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$   \\
$p_1$ : $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$ \\
\hline
\end{tabular}
\end{table}

%\tikzmark{p1} 
%\begin{tikzpicture}[remember picture, overlay]
%  \path[draw=red,thick,<->]<4-> (pic cs: p0) to [] (pic cs: p1);
%\end{tikzpicture}

\hspace{-20pt}
\begin{minipage}{.25\textwidth}
\small
\onslide<2-> 
\begin{table}
\caption*{SC}
\begin{tabular}{ >{\columncolor{ltgrey}}l >{\columncolor{ltgrey}}r}
$\b{p_0}, \b{p_1}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$ \\
$\r{q_0}, \r{q_1}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$
\onslide<3-> 
\\
$\b{p_0}, \r{q_0}, \b{p_1}, \r{q_1}$ & $(\b{1}; \r{1})$ \\
$\b{p_0}, \r{q_0}, \r{q_1}, \b{p_1}$ & $(\b{1}; \r{1})$ \\
$\r{q_0}, \b{p_0}, \b{p_1}, \r{q_1}$ & $(\b{1}; \r{1})$ \\
$\r{q_0}, \b{p_0}, \r{q_1}, \b{p_1}$ & $(\b{1}; \r{1})$ \\
\end{tabular}
\end{table}
\end{minipage}
\pause
%
\hspace{5pt}
\begin{minipage}{.6\textwidth}
\small
\onslide<4->
%\vspace{-1em}
%"delay the effect of a store past any load from a different location."
%\vspace{-11pt}
%\vspace{-2pt}
\begin{table}
\caption*{TSO}
\begin{tabular}{ | l r | l r | l r }
$\underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \b{p_1}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$ \\
$\r{q_0}, \r{q_1}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$ \\
$\underline{\b{p_1}}, \r{q_0}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \b{p_1}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & $(\b{0}; \r{0})$ \\
$\underline{\b{p_1}}, \r{q_0}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{0}; \r{0})$  &  $\b{p_0}, \underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & $(\b{0}; \r{0})$ \\
$\r{q_0}, \underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{1}; \r{1})$  &  $\underline{\r{q_1}}, \b{p_0}, \b{p_1}, \underline{\r{q_0}}$ & $(\b{0}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & $(\b{0}; \r{0})$ \\
$\r{q_0}, \underline{\b{p_1}}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \b{p_0}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & $(\b{0}; \r{0})$ \\
\end{tabular}
\end{table}
\end{minipage}

@@

\begin{frame}{Verification of concurrent software} {Example: Store buffering (hardware relaxations)}
\begin{center}
\scalebox{0.7}{
\ttfamily
\begin{tabular}{ |>{\color{dkblue}}c | >{\color{dkred}}c| }
\hline
\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P & Q \\ \hline
$p_0$ : $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$   \\
$p_1$ : $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$ \\
\hline
\end{tabular}
}
\end{center}

\begin{figure}
\includegraphics[scale=0.37]{img/x86-arch-full.png}
\caption{An x86-TSO abstract machine~\cite{sewell2010x86}}
%\label{fig:x86-arch}
\end{figure}
\end{frame}

\begin{frame}{The weak memory model} {Axiomatic semantics: The definition}

@@

\begin{frame}{The weak memory model} {Testing the candidate executions}
\vspace{10pt}
\begin{minipage}{.3\linewidth}
\hfill
\end{minipage}
%
\noindent\begin{minipage}{.4\linewidth}
\scalebox{0.7}{
\ttfamily
\begin{tabular}{ |>{\color{dkblue}}c | >{\color{dkred}}c| }
\hline
\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P & Q \\ \hline
$p_0$ : $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$   \\
$p_1$ : $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$ \\
\hline
\end{tabular}
}
\end{minipage}
%
\noindent\begin{minipage}{.25\linewidth}
\end{minipage}
\vspace{-5pt}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{img/candidate.png}
\caption{The four candidate executions allowed under x86-TSO}
\end{figure}
\end{frame}

\begin{frame}{The weak memory model} {Testing the candidate executions}
\vspace{10pt}
\begin{minipage}{.3\linewidth}
\hfill
\end{minipage}
%
\noindent\begin{minipage}{.4\linewidth}
\scalebox{0.7}{
\ttfamily
\begin{tabular}{ |>{\color{dkblue}}c | >{\color{dkred}}c| }
\hline
\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P & Q \\ \hline
$p_0$ : $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$   \\
$p_1$ : $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$ \\
\hline
\end{tabular}
}
\end{minipage}
%
\raggedleft{
\noindent\begin{minipage}{.25\linewidth}
\scriptsize
SC model: \\
... \\
\texttt{fr = (rf\^\,-1; co)} \\
\color{red}\texttt{acyclic(fr $\cup$ po)}
\end{minipage}
}
\vspace{-5pt}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{img/candidate-red.png}
\caption{The four candidate executions allowed under x86-TSO}
\end{figure}

@@

\begin{frame}{Tools for memory model-aware analysis}
\begin{itemize}
\item \texttt{diy} tool suite:
  \begin{itemize}
  \item \tool{diy}, \tool{diycross} and \tool{diyone}, litmus tests generators,
  \item \tool{litmus}, a litmus test concrete executor, and 
  \item \tool{herd}, a weak memory model simulator;
  \end{itemize}
%\item the stateless model checkers (\tool{CHESS}~\cite{musuvathi2008fair}, \tool{Nidhugg}~\cite{abdulla2017stateless}); %TODO: description of statelessses
\item the stateless model checkers (\tool{CHESS}, \tool{Nidhugg}); %TODO: description of statelessses
\item the tool for automated synthesis of the synchronisation primitives \tool{musketeer};
\item the instrumenting compiler \tool{goto-cc} which is a part of \tool{CBMC} model checker;
\item the tool \texttt{Porthos} for analysing the portability of the C programs;
\item and others.
\end{itemize}
\end{frame}

\subsection{Portability Analysis}

\begin{frame}{Portability analysis} {The \texttt{Porthos} tool}

\begin{itemize}
\item Let the function $\textit{cons}_{\mathcal{M}}(P)$ calculate the set of executions of program $P$ consistent under the memory model $\mathcal{M}$.
\end{itemize}

\begin{definition}[Portability~\cite{Porthos17a}]
Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models.
The program $P$ is portable from $\mathcal{M_S}$ to $\mathcal{M_T}$ if 
$\textit{cons}_{\mathcal{M_T}}(P) \subseteq \textit{cons}_{\mathcal{M_S}}(P)$
\end{definition}

\begin{itemize}
\item Portability as an SMT-based bounded reachability problem: \\
${\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}}$
\item $\texttt{SAT}(\phi) \implies $ the portability bug
\end{itemize}

%end{tabular}
\end{frame}

\subsection{SMT-encoding}

%\begin{frame}{Encoding for the control-flow}% {The \texttt{Porthos v1} encoding scheme}
%\begin{itemize}
%\item In \texttt{Porthos}, the high-level instructions were represented in the SMT-formula by separate variables.
%\item For example, the sequential instruction $i_1 = i_2; i_3$ was encoded as
%$\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
%\end{itemize}
%\end{frame}

>>>>

\documentclass[aspectratio=149]{beamer}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian,english]{babel}

\usepackage{outlines}

\definecolor{ltgrey}{rgb}{0.85,0.85,0.85}
\definecolor{ltblue}{RGB}{101,147,245}
\definecolor{dkblue}{rgb} {0,   0.1, 0.6}
\definecolor{dkred} {rgb} {0.6, 0.1, 0}
\renewcommand{\r}[1]{\ensuremath{\textcolor{dkred}{#1}}}
\renewcommand{\b}[1]{\ensuremath{\textcolor{dkblue}{#1}}}

\usepackage{xparse}
\usepackage{tikz}
\usetikzlibrary{tikzmark}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.pathmorphing,decorations.markings}
\usetikzlibrary{shapes}
\usetikzlibrary{patterns}
\usetikzlibrary{automata}

@@

\newcommand*{\yellowemph}[1]{%
  \tikz[baseline=(X.base)] \node[rectangle, fill=yellow, inner sep=0mm] (X) {#1};%
}

\newcommand*{\rus}[1]{%
  %\textcolor{black}{
  \textit{#1}
  %}
}

@@

%\usepackage{../inc/listingset}
\usepackage{listings}
\lstset{escapeinside={[*}{*]}}

@@

\usepackage[backend=biber]{biblatex}
\addbibresource{../bib/ms-thesis.bib}

\usepackage{array}
%\usepackage{xcolor}

@@

% There are many different themes available for Beamer. A comprehensive
% list with examples is given here:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
% You can uncomment the themes below if you would like to use a different
% one:
%\usetheme[school={SCI}]{Aalto}
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{default}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}\colorbox{yellow}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier} %!
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

%\beamertemplatenavigationsymbolsempty  %no navigation bar
%\setbeamertemplate{footline}[frame number]
%\setbeamertemplate{headline}{}

\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber%/\inserttotalframenumber
}

%\setbeamercovered{transparent}
%\setbeamertemplate{footline}{}

%\setbeamertemplate{page number in head/foot}[framenumber]

%\setbeamertemplate{frametitle}[default][left]

%\titlegraphic{\includegraphics[height=1cm,width=2cm]{logo1}}
%\titlegraphicii{\includegraphics[height=1cm,width=2cm]{logo2}}
\title{\nohyphens{\textbf{Automated~Analysis~of~Weak~Memory~Models}}}

@@

\institute
{
  \inst{1}%
  \textbf{Aalto University} (Espoo, Finland) \\
  Department of Computer Science, \\
  School of Science
  \and
  \inst{2}%
  \textbf{ITMO University} (Saint Petersburg, Russia) \\
  Faculty of Information Security \\
  and Computer Technologies
}

@@

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
%\AtBeginSection[]
%{
%  \begin{frame}<beamer>{Outline}
%    \tableofcontents[currentsection]%,currentsubsection]
%  \end{frame}
%}

% Let's get started
\begin{document}
%\beamertemplatenavigationsymbolsempty %remove navigation bar

@@

%\begin{frame}{Outline}
%  \tableofcontents
%\end{frame}

%\section{Introduction}
%\subsection{The problem}

\begin{frame}{Task statement \rus{(Цель работы)}}
\begin{itemize}
\item Improve the proof-of-concept memory model-aware analysis tool \texttt{Porthos}~\cite{Porthos17a} by extending support for the input language.
\end{itemize}
\end{frame}

\begin{frame}{Task specification \rus{(Задачи работы)}}
\begin{itemize}
\item Study the general framework for memory model-aware analysis of concurrent programs~\cite{alglave2010shared};
\item Review existing tools for memory model-aware analysis;
%\item Examine the architecture of \texttt{Porthos}, its strengths and weaknesses;
%\item Design a new architecture for \texttt{PorthosC} that allow to extend the input language to the (large subset of) C language, be robust, transparent, efficient and extensible.
\item Develop a \textit{C compiler infrastructure} as a part of an abstract interpretation engine for the new tool \tool{PorthosC};
\item Improve the SMT-encoding scheme for an \textit{arbitrary control-flow};
\item Enhance performance, extensibility, reliability and maintainability of the tool.
\end{itemize}
\end{frame}

%\subsection{The background: weak memory model-aware analysis}

\begin{frame}{Verification of concurrent software} {Motivating example: Write-write reordering (compiler  relaxations)}
%1) mckenney  http://www.rdrop.com/users/paulmck/scalability/paper/whymb.2010.06.07c.pdf
%2) http://preshing.com/20120515/memory-reordering-caught-in-the-act/
\begin{table}
\centering
\ttfamily
\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
\hline
\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P & Q \\ \hline
\tikzmark{p0}$p_0$: $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$\tikzmark{q0} \\
\tikzmark{p1}$p_1$: $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$\tikzmark{q1} \\
\hline
\multicolumn{2}{|c|}{ exists ($\b{r_p}=0 \land \r{r_q}=0$) } \\
\hline
\end{tabular}
\end{table}

\onslide<2->
\begin{minipage}{.25\textwidth}
\small
\begin{table}
\only<1>{\caption*{~\\~}}
\only<2>{\caption*{Single thread\\(no concurrency)}}
\only<3>{\caption*{Sequential Consistency\\(classical concurrency)}}
\only<4->{\caption*{~\\~}}
\only<4->{
  \begin{tikzpicture}[overlay, remember picture]
    \draw [<->,thick,red,looseness=3] ([xshift=-5pt,yshift=3pt]{pic cs:p0}) to[bend right=90] ([xshift=-5pt]{pic cs:p1});
    \draw [<->,thick,red,looseness=3] ([xshift=8pt,yshift=3pt]{pic cs:q0}) to[bend left=90] ([xshift=5pt]{pic cs:q1});
  \end{tikzpicture}%
}
%"A multiprocessor is sequentially consistent if the result of any
%execution is the same as some sequential order, and within any
%processor, the operations are executed in program order"
%– Lamport, 1979
\begin{tabular}{ l r }
$\b{p_0}, \b{p_1}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$ \\
$\r{q_0}, \r{q_1}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$
\onslide<3->
\\
$\b{p_0}, \r{q_0}, \b{p_1}, \r{q_1}$ & $(\b{1}; \r{1})$ \\
$\b{p_0}, \r{q_0}, \r{q_1}, \b{p_1}$ & $(\b{1}; \r{1})$ \\
$\r{q_0}, \b{p_0}, \b{p_1}, \r{q_1}$ & $(\b{1}; \r{1})$ \\
$\r{q_0}, \b{p_0}, \r{q_1}, \b{p_1}$ & $(\b{1}; \r{1})$ \\
\end{tabular}
\end{table}
\end{minipage}
%
\onslide<4->
\begin{minipage}{.6\textwidth}
\small
%"delay the effect of a store past any load from a different location."
\begin{table}
%http://events17.linuxfoundation.org/sites/events/files/slides/dbueso-elc2016-membarriers-final.pdf
%SPARC, x86 (Intel, AMD)
%Similar to SC, but:
%‒ Loads may be reordered with writes.
%Relaxed Models
% Arbitrary reorder limited only by explicit memorybarrier
%instructions.
\caption*{Total Store Order\\(e.g., x86)}
\begin{tabular}{ | l r | l r | l r } % >{\columncolor{ltgrey}}
$\underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \b{p_1}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$ \\
$\r{q_0}, \r{q_1}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$ \\
$\underline{\b{p_1}}, \r{q_0}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \b{p_1}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}} \\
$\underline{\b{p_1}}, \r{q_0}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}} \\
$\r{q_0}, \underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{1}; \r{1})$  &  $\underline{\r{q_1}}, \b{p_0}, \b{p_1}, \underline{\r{q_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}}  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}} \\
$\r{q_0}, \underline{\b{p_1}}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \b{p_0}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & {\only<5>\yellowemph{$(\b{0}; \r{0})$}} \\
\end{tabular}
\end{table}
\end{minipage}

@@

\begin{frame}{Verification of concurrent software} {Motivating example: Store buffering (hardware relaxations)}
\begin{center}
\scalebox{0.75}{
\ttfamily
\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
\hline
\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P & Q \\ \hline
$p_0$: $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$\\
$p_1$: $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$\\
\hline
\multicolumn{2}{|c|}{ exists ($\b{r_p}=0 \land \r{r_q}=0$) } \\
\hline
\end{tabular}
}
\end{center}

\begin{figure}
\includegraphics[scale=0.37]{img/x86-arch-full.png}
\caption{An x86-TSO abstract machine~\cite{sewell2010x86}}
\only<2->{
    \begin{tikzpicture}[overlay]
    \node at (0.1, 3.1) {\fcolorbox{yellow}{yellow}{
        \includegraphics[scale=0.2]{img/table.png}} };
    \end{tikzpicture}
}
\end{figure}

\begin{frame}{Weak memory model-aware analysis} {Event-based program representation}

@@

\begin{frame}{Weak memory model-aware analysis} {Testing candidate executions}
\vspace{10pt}
\begin{minipage}{.33\linewidth}
\hfill
\end{minipage}
%
\noindent\begin{minipage}{.39\linewidth}
\scalebox{0.65}{
\ttfamily
\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
\hline
\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P & Q \\ \hline
$p_0$: $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$\\
$p_1$: $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$\\
\hline
\multicolumn{2}{|c|}{ exists ($\b{r_p}=0 \land \r{r_q}=0$) } \\
\hline
\end{tabular}
}
\end{minipage}
%
\noindent\begin{minipage}{.23\linewidth}
\only<2->{
\scriptsize
SC model: \\
... \\
\texttt{fr = (rf\^\,-1; co)} \\
\color{red}\texttt{acyclic(fr $\cup$ po)}
}
\end{minipage}
\vspace{-5pt}
\only<1>{
\begin{figure}
%\centering
\includegraphics[width=\linewidth]{img/candidate.png}
\caption{The four candidate executions allowed under x86-TSO}
\end{figure}
}
\only<2->{
\begin{figure}
%\centering
\includegraphics[width=\linewidth]{img/candidate-red.png}
\caption{The four candidate executions allowed under x86-TSO}
\end{figure}
}
\only<3>{
  \begin{tikzpicture}[overlay]
  \node at (6.1, 4.1) {\fcolorbox{yellow}{yellow}{
    \includegraphics[scale=0.11]{img/non-portable.png}} };
  \end{tikzpicture}
}

@@

\begin{frame}<presentation:0>{Tools for memory model-aware analysis}
\begin{itemize}
\item \texttt{diy} tool suite:
  \begin{itemize}
  \item \tool{diy}, \tool{diycross} and \tool{diyone}, litmus tests generators,
  \item \tool{litmus}, a litmus test concrete executor, and 
  \item \tool{herd}, a weak memory model simulator;
  \end{itemize}
%\item the stateless model checkers (\tool{CHESS}~\cite{musuvathi2008fair}, \tool{Nidhugg}~\cite{abdulla2017stateless}); %TODO: description of statelessses
\item the stateless model checkers (\tool{CHESS}, \tool{Nidhugg}); %TODO: description of statelessses
\item the tool for automated synthesis of the synchronisation primitives \tool{musketeer};
\item the instrumenting compiler \tool{goto-cc} which is a part of \tool{CBMC} model checker;
\item the tool \texttt{Porthos} for analysing the portability of the C programs;
\item and others.
\end{itemize}
\end{frame}

\begin{frame}{Portability analysis} {The \tool{Porthos} tool}
\begin{minipage}{.49\textwidth}
\only<1->{
  \begin{figure}
  \includegraphics[height=0.4\textheight,keepaspectratio]{img/the-problem.png}
  \caption{The illustration of the portability problem~\cite{Porthos17slides}}
  \end{figure}
}
\end{minipage}
%
\begin{minipage}{.5\textwidth}
\only<2->{
  \begin{itemize}
    \begin{definition}[Portability~\cite{Porthos17a}]
    %SAY: Let $\textit{cons}_{\mathcal{M}}(P)$ be a function that calculates the set of executions of the program $P$ that are consistent under the memory model $\mathcal{M}$
    %Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models.
    The program $P$ is portable from the source memory model $\mathcal{M_S}$ to the target one $\mathcal{M_T}$ if
    $\textit{cons}_{\mathcal{M_T}}(P) \subseteq \textit{cons}_{\mathcal{M_S}}(P)$
    \end{definition}
    \item Portability as an SMT-based bounded reachability problem: \\
    $\phi = $\only<3>\yellowemph{$\phi_{CF}$}$\land $\only<3>\yellowemph{$\phi_{DF}$}$\land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$
    \item $\texttt{SAT}(\phi) \Longrightarrow $ portability bug
  \end{itemize}
}
\end{minipage}
\end{frame}

\begin{frame}{\tool{PorthosC}: Architecture}{Main components}
\vspace{-30pt}
\begin{minipage}{.49\textwidth}
\begin{itemize}
\item The enhanced version of \tool{Porthos}, that is able to analyse C programs, has got a new name \textit{PorthosC}.
\end{itemize}
\hfill
\end{minipage}
%
\begin{minipage}{.5\textwidth}
\only<1>{\begin{figure}
  \includegraphics[height=.98\textheight,keepaspectratio]{img/arch/general_arch-no_numbering.png}
\end{figure}}
\only<2>{\begin{figure}
  \includegraphics[height=.98\textheight,keepaspectratio]{img/arch/general_arch-no_numbering-labels.png}
\end{figure}}
\only<3>{\begin{figure}
  \includegraphics[height=.98\textheight,keepaspectratio]{img/arch/general_arch-no_numbering-visitors.png}
\end{figure}}
\only<4>{\begin{figure}
  \includegraphics[height=.98\textheight,keepaspectratio]{img/arch/general_arch-no_numbering-compiler.png}
\end{figure}}
\only<5>{\begin{figure}
  \includegraphics[height=.98\textheight,keepaspectratio]{img/arch/general_arch-no_numbering-smt.png}
\end{figure}}
\end{minipage}
\end{frame}

\begin{frame}{\tool{PorthosC}: Architecture}{Abstract interpretation module}
  \vspace{-40pt}
\begin{minipage}{.4\textwidth}
\begin{itemize}
\item The abstract interpretation module is an essential part of the compiler of C code into the event-flow graph representation (\texttt{X-graph}).
\end{itemize}
\hfill
\end{minipage}
%
\begin{minipage}{.5\textwidth}
\only<1>{\begin{figure}
  \scalebox{0.9}{\includegraphics[height=.99\paperheight]{img/compiler/X-compiler.png}}
\end{figure}}
\only<2>{\begin{figure}
  \scalebox{0.9}{\includegraphics[height=.99\paperheight]{img/compiler/X-compiler-1.png}}
\end{figure}}
\only<3>{\begin{figure}
  \scalebox{0.9}{\includegraphics[height=.99\paperheight]{img/compiler/X-compiler-2.png}}
\end{figure}}
\only<4>{\begin{figure}
  \scalebox{0.9}{\includegraphics[height=.99\paperheight]{img/compiler/X-compiler-3.png}}
\end{figure}}
\only<5>{\begin{figure}
  \scalebox{0.9}{\includegraphics[height=.99\paperheight]{img/compiler/X-compiler-4.png}}
\end{figure}}
\end{minipage}
\end{frame}
